---
title: "4a - Aniket Maheshwari"
output: word_document
---


Setting up our environment and importing important libraries:
```{r}
### Clear the environment 
rm(list = ls())


### First we will set the directory of the R script 
setwd("C:/Users/anike/Desktop/Sem 1/EAS 506 Statistical Data Mining/Homework/Homework 4")


## Loading all the libraries 
library(ISLR)
library(corrplot)
library(MASS)
library(klaR)
library(leaps)
library(lattice)
library(ggplot2)
library(corrplot)
library(car)
library(caret)
library(class)
library(boot)
#install.packages("bootstrap")
library(bootstrap)
```


Importing the dataset: 

```{r}
data("Boston")
dim(Boston)
str(Boston)
data1 <- Boston
summary(Boston)
```

So the dataset 'Boston' has 506 rows and 14 columns or features. All the features are numerical variables.

Before starting EDA, first i’ll check whether the data has any missing values or not:

```{r}
NAmat = matrix(as.numeric(is.na(data1)) , ncol = 14)
nonNAdx = which(rowSums(NAmat) == 0)
length(nonNAdx)
dim(data1)
```

so there are no missing value as length of nonNAdx is equal to number of rows in dataset Boston.


Now, the value of the different feature are in different size of scale. For example, crim is between 0.00632 - 88.97620 range whereas tax is in between 187-711 range. So i need to normalize this dataset so that all the features are in one scale before working on the dataset.

Normalization: 
```{r}


normalize <- function(x) {
  (x -min(x)) / (max(x) - min(x))
  
}

Boston_Norm <- as.data.frame(lapply(data1[1:14], normalize))
head(Boston_Norm , 1)

```


Now the full data set is normalized.

Splitting the data set into test and train data:
Now, I'll split the data into test and train dataset in 2:8 ratio. After the splitting, the train dataset will have 404 rows and 14 columns and test data set will have 102 rows and 14 columns.
```{r}
set.seed(1)
sample_size <- floor(0.80 * nrow(Boston_Norm))
set.seed(1)
train_indexes <- sample(seq_len(nrow(Boston_Norm)), size = sample_size)
training_data <- Boston_Norm[train_indexes ,]
testing_data <- Boston_Norm[-train_indexes ,]
dim(training_data)
dim(testing_data)
```

Best Subset Selection: 
Best subset selection is a method that aims to find the subset of independent variables (Xi) that best predict the outcome (Y) and it does so by considering all possible combinations of independent variables.

The R function regsubsets() [leaps package] can be used to identify different best models of different sizes.

```{r}
boston_best_subset <- regsubsets(medv ~ . , data = training_data , nbest = 1,  really.big = TRUE , nvmax = 13) 
boston_best_subset_summary <- summary(boston_best_subset)
boston_best_subset_summary
```


Model Selection using best subset regression:
The summary() function has some metrics that allows us to determine the best overall model with least error value. 
I'll use AIC/CP and BIC metric.

1. AIC/CP
```{r}
which.min(boston_best_subset_summary$cp)
# model with 11 features has the lowest AIC error 
x11()
plot(boston_best_subset_summary$cp ,xlab=" Number of Variables ",ylab=" Cp", type="l")
points(11, boston_best_subset_summary$cp[11],col="red" , cex = 2 , pch = 20)
```



```{r}
x11()
plot(boston_best_subset, scale = "Cp")
```

So AIC gives minimum error in model with 11 variables.


2. BIC 
```{r}
which.min(boston_best_subset_summary$bic)
# model with 9 features has the lowest BIC error 
x11()
plot(boston_best_subset_summary$bic ,xlab=" Number of Variables ",ylab=" BIC", type="l")
points(9, boston_best_subset_summary$bic[9],col="red" , cex = 2 , pch = 20)
```


```{r}
x11()
plot(boston_best_subset, scale = "bic")
```

So BIC gives minimum error in model with 9 variables.

3. K-Fold Cross Validation: 
The k-fold Cross-validation consists of first dividing the data into k subsets. Each subset[10%] is used as test dataset and remaining[90%] subset serves as training dataset. 

a) 5-fold cross validation:

```{r}
predict.regsubsets = function(object,newdata,id , ...){
  form = as.formula((object$call[[2]]))
  mat = model.matrix(form, newdata)
  coefi = coef(object , id = id)
  xvars = names(coefi)
  mat[,xvars]%*%coefi
  
}


set.seed(1)
k = 5

five_fold_error <- matrix(NA, k, 13)

five_fold = sample(1:k, nrow(training_data), replace = TRUE)


for (i in 1:k)
{
  best_subset = regsubsets(medv~., data = training_data[five_fold!=i, ], nvmax = 13)
  
  for (j in 1:13)
  {
    predictions = predict(best_subset, training_data[five_fold==i, ], id=j)
    five_fold_error[i,j] = mean((training_data$medv[five_fold==i]-predictions)^2)
  }
  
}

five_fold_error
```

5 - fold cross validation error: 
```{r}
mse.models <- apply(five_fold_error, 2, mean)  
mse.models
```


Plotting the 5-fold cross-validation error: 
```{r}
x11()
plot(mse.models , pch=19, type="b",xlab="Predictors",ylab="MSE errors")
which.min(mse.models)
```
5 - fold cross validation has 11 variable model with the lowest error. 

b) 10-fold cross validation:

```{r}
set.seed(1)
k = 10

ten_fold = sample(1:k, nrow(training_data), replace = TRUE)


ten_fold_cross = matrix(NA, k, 13)

for (i in 1:k)
{
  best_subset1 = regsubsets(medv~., data = training_data[ten_fold!=i, ], nvmax = 13)
  
  for (j in 1:13)
  {
    predictions1 = predict(best_subset1, training_data[ten_fold==i, ], id=j)
    ten_fold_cross[i,j] = mean((training_data$medv[ten_fold==i]-predictions1)^2)
  }
  
}
ten_fold_cross
```

10-Fold cross-validation error:
```{r}
mse.models1 = apply(ten_fold_cross, 2, mean)
which.min(mse.models1)
```
Plotting the errors:
```{r}
x11()
plot(mse.models1 , pch=19, type="b",xlab="Predictors",ylab="MSE errors")

```

Like 5 fold cross-validation, 10 fold cross validation also gives 11 variable model with lowest error rate. 



4. Bootstrap .632 estimator: 
Now, I'll use bootstrap estimator for model selection. I will use bootstrap .632 estimator for prediction error.

```{r}

beta.fit <- function(X,Y)
{
  lsfit(X,Y)
}

beta.predict<- function(fit, X)
{
  cbind(1,X)%*%fit$coef
}

sq.error<- function(Y, Yhat)
{
  (Y-Yhat)^2
}


# Creating X and Y 

X <- Boston_Norm [,1:13]
Y <- Boston_Norm [,14]
dim(X)
length(Y)

select <- boston_best_subset_summary$outmat

error_values = c()
 
for (i in 1:13){
  temp <- which(select[i,] == "*")
  res<- bootpred(X[,temp], Y, nboot = 50, theta.fit = beta.fit, theta.predict = beta.predict,  err.meas = sq.error)
  error_values <- c(error_values, res[[3]])
}

error_values


```
Plotting the Bootstrap .632 prediction error: 
```{r}
x11()
plot(error_values , pch=19, type="b",xlab="Predictors",ylab="Bootstrap .632 Error Values")

```

So, Bootstrap estimator .632 gives 12 variable model with least error rate. 


Comparing all the model selection methods: 

```{r}
output <- data.frame(
  AIC = which.min(boston_best_subset_summary$cp),
  BIC = which.min(boston_best_subset_summary$bic),
  Five_Fold_Cross_Validation = which.min(mse.models),
  Ten_Fold_Cross_Validation = which.min(mse.models1),
  Bootstrap_632 = which.min(error_values)
)

output

```





